Resonant Scattering in a Uniform Sphere with Large Optical Depth

### 2021-06-22 ms.tex
Replacing \nu with \sigma in function declarations
Proof reading up to Monte Carlo section

The solution of the radiative transfer equation for resonant scattering of Lyman $\alpha$ photons for a uniform sphere of constant gas density is found in the limit of large line center optical depth. A monochromatic source of photons is assumed at the center of the sphere, and the intensity within the sphere and at the surface is found with the Eddington approximation for the angular dependence. The solution can be represented a sum of three terms: a simple, analytic solution of the inhomogeneous equation which diverges at the center of the sphere and at the emission frequency; a semi-analytic solution of the homogeneous equation which is required for the mean intensity $J=0$ at the surface; and finally, a solution of the homogeneous equation which enforces the correct, zero ingoing intensity boundary conditions at the surface. It is the compact, analytic form as well as the piece accounting for the correct, frequency dependent boundary condition that is novel in this study. The analytic solution is compared to numerical solutions with the Monte-Carlo method, which are valid at arbitrary optical depth, and the deviations from exact solution are investigated.

@@

The outer layers of planetary atmospheres are central to a planet's evolution, since they can shelter the lower atmosphere from high energy radiation and regulate the escape of gas into space. In atmospheres with thick layers of atomic hydrogen, stellar Lyman Alpha (Ly$\alpha$) radiation is especially important in this process as the optical depth at line center is large. Ionization and subsequent recombination can generate \lya within the planet's atmosphere, dissociating molecules in lower layers and causing heating by collisional de-excitation. It can also excite H atoms to the 2p state, creating a population of Balmer-line absorbers that can be observed via transmission spectroscopy. \todo{Need citations for these statements?}

@@

Hubble Space Telescope (HST) observations with STIS have found large Lyman $\alpha$ transit depths, indicating an extended distribution of hydrogen around a handful of exoplanets: \citet{2003Natur.422..143V, 2012A&A...543L...4L, 2012A&A...547A..18E, 2015Natur.522..459E,  2017A&A...597A..26B, 2017A&A...599L...3B, 2017A&A...602A.106B, 2018A&A...620A.147B, 2019AJ....158...50W, 2019EPSC...13.1928L, 2020ApJ...888L..21G,2021arXiv210309864B}. These observations have revealed a population of atoms extending out to distances of order the planetary radius for several planets around bright, nearby stars, motivating a study of the physics of \lya interactions with H atoms surrounding a planet. The importance of the \lya line in the atmospheric dynamics of these systems necessitates a careful treatment of resonant scattering in order to construct accurate models of atmospheric heating and escape. However, the extreme optical depths at \lya line center introduce a computational barrier to simulating radiative transfer via Monte Carlo methods.

@@

A third class of methods are modified random walk methods \citep{1984JCoPh..54..508F,2009A&A...497..155M,Robitaille2010}. In this case a spherical volume (usually the largest within a simulation cell) is chosen when the scattering mean-free-path is small compared with the zone size.  An outgoing photon is then randomly sampled on the surface of this outgoing sphere by drawing its properties from distributions in outgoing frequencies, directions and escape times, based on solutions to the diffusion equation within a uniform sphere. A method similar to this has been applied by \citet{2006ApJ...645..792T} to Lyman $\alpha$ transfer using the \cite{1990ApJ...350..216N} solution.

\todo{FOR REVIEW: a summative few sentences connecting the MC methods to what we're doing in this paper:}

\proposed{The need for accurate distributions from which to draw outgoing photon properties motivates this work; namely, an analytic solution to the steady-state problem of \lya resonant scattering in spherical geometry and a semi-analytic solution to time-dependent \lya diffusion through the same medium. These will be shown to accurately represent the photon distributions in frequency and time-to-escape at the surface of a spherical volume of gas, highlighting their potential for use as a Monte Carlo accelerant in magnetohydrodynamic calculations of phenomena involving \lya resonance-line radiation.}

@@

\citet{1973MNRAS.162...43H, 1974MNRAS.166..373H} demonstrated that resonance-line radiation transfer in low-density, optically-thick media can be described by the Poisson equation. However, the accuracy of their solution is limited by the treatment of the boundary conditions. In their eigenfunction expansion, the spatial and frequency components of the differential equation are separated, yet the spatial eigenvalues are defined as being explicitly frequency-dependent. The resulting solution for large optical depths would suggest the mean intensity at the boundary of the slab is zero, which is unphysical. A more detailed discussion of the boundary condition is included in Section \ref{sec:no_destruction}. \citet{1990ApJ...350..216N} follows this approach in an attempt to extend the result to media of intermediate optical depth, including the effects of scattering in the Doppler core of the line, but again does not use the correct boundary condition. \citet{?} improved on this approach, producing a divergent, inhomogeneous solution to the problem in 1D slab geometry that includes the effect of photon absorption within the medium. \citet{2006ApJ...649...14D} generalize the problem to spherical geometry, which is the same physical set up used here. Following this, \citet{2020arXiv200509692L} presents closed-form solutions for the radiation field in spherical geometry, utilizing the ``core skipping'' method discussed in \citet{2015MNRAS.449.4336S} to handle photon frequency redistribution within the Doppler core. This technique involves drawing the atom's velocity from a truncated Gaussian distribution, forcing photons back into the wing when they scatter into the core. This is motivated by the assumption that all core scattering results in zero spatial diffusion until the photon is redistributed back out into the wing where the mean free path becomes much larger and it can diffuse in space again. However, it is not clear whether this prescription is fully independent of the choice of velocity distribution.

@@

\subsection{Derivation of the Steady-State Solution}
Now, consider a sphere of radius $R$ and uniform density, with line-center optical depth $\tau_0 = (k/\sqrt{\pi}\Delta)R$. Here, the Doppler width is $\Delta = \nu_0 v_{\rm th}/c$, where $v_{\rm th}=\sqrt{2k_{\rm B}T/m_{\rm H}}$ is the thermal speed of hydrogen atoms of mass $m_{\rm H}$ and temperature $T$, and $c$ is the speed of light. After changing the photon frequency variable in Eq. \ref{eq:rteqn} and performing all angular integrals, we obtain Eq. \ref{eq:finaleqn} (see Appendix \ref{app:rteqn_derivation}) and set $p=\alpha_{\rm abs}=0$ to eliminate photon destruction terms. With a photon emission term given by Eq. \ref{eq:jem}, this gives the transfer equation
\be
\nabla^2 J + \left( \frac{k}{\Delta} \right)^2 \frac{\partial^2 J}{\partial \sigma^2} & = & 
- \frac{ \sqrt{6} kL}{4\pi \Delta^2} \delta^3(\vec{x} - \vec{x}_s) \delta (\sigma - \sigma_{\rm s})
\label{eq:rt_no_destr}
\ee
where the form of the source term is in Eq. \ref{eq:jem_v2} and a change of variables in frequency has been performed using $d\sigma/dx = \sqrt{2/3}\ \phi(x)^{-1}$. The boundary condition is that there is no incoming flux at the surface, which can be written
\citep{1986rpa..book.....R}
\be
J & = & \sqrt{3} H
\label{eq:bc}
\ee
at $r=R$.

First consider the divergent solution. In terms of the dimensionless variables $x_1 = k(x-x_{\rm s})/\Delta$, $x_2=k(y-y_{\rm s})/\Delta$, $x_3=k(z-z_{\rm s})/\Delta$ and $x_4=\sigma - \sigma_{\rm s}$, this equation can be written as a Laplacian in the 4-dimensional space as
\be
\nabla_4^2 J & = & - \frac{\sqrt{6} k^2 L}{4\pi \Delta^3} \delta^4(x).
\ee
In terms of the 4-dimensional radius $r_4 = \sqrt{ (k/\Delta)^2\left[ (x-x_{\rm s})^2 + (y-y_{\rm s})^2 + (z-z_{\rm s})^2 \right] + (\sigma - \sigma_{\rm s})^2 }$, and assuming spherical symmetry about the source point, the equation becomes
\be
\frac{1}{r_4^3} \frac{\partial}{\partial r_4} \left( r_4^3 \frac{\partial J}{\partial r_4} \right) & = & - \frac{\sqrt{6} k^2 L}{4\pi \Delta^3} 
\frac{\delta(r_4)}{2\pi^2 r_4^3},
\ee
where the forms of the 4-dimensional Laplacian and delta function have been inserted. The ``divergent" solution due to the delta function is then
\be
J_{\rm d} & = & 
\frac{(\sqrt{6}k^2L)/(16\pi^3 \Delta^3)}{ (k/\Delta)^2\left[ (x-x_{\rm s})^2 + (y-y_{\rm s})^2 + (z-z_{\rm s})^2 \right] + (\sigma - \sigma_{\rm s})^2}.
\label{eq:Jd}
\ee
While this solution does not satisfy the surface boundary condition (Eq. \ref{eq:bc}), it is accurate near the delta-function source.
This solution generalizes the particular solution for slab geometry found by \citet{1990ApJ...350..216N}. Henceforth we will focus on the source at the center of the sphere, although we will retain the emission frequency at $\sigma=\sigma_{\rm s}$. The radial flux corresponding to Eq. \ref{eq:Jd} is
\be
H_{\rm d} & = & - \frac{1}{3k\phi} \frac{\partial J_d}{\partial r}
=  \left( \frac{1}{3k\phi} \right) 
\left( \frac{ \sqrt{6}k^3L }{ 8\pi^3 \Delta^4} \right)
\left( \frac{k r/\Delta}{ \left[ (kr/\Delta)^2 + (\sigma-\sigma_{\rm s})^2 \right]^2 } \right).
\label{eq:Hd}
\ee
It will be shown that Eq. \ref{eq:Hd} is a good approximation at $r=R$ near line center. However Eq. \ref{eq:Jd} is not a good approximation to the true solution, as it is too large at $r=R$ by a factor of $J_{\rm d}(R,\sigma)/ H_{\rm d}(R,\sigma) \sim a\tau_0/x^2 \sim (a\tau_0)^{1/3} \gg 1$.

@@

In the limit of large optical depth $B$, \citealt{1973MNRAS.162...43H} approximates the eigenvalues as $\lambda_n B \simeq \pi (n-1/2)$, which gives $J=0$ at the surface. This is essentially the result in Eq. \ref{eq:J0} above. However, since this solution gives zero at the surface, their Eq. 34 subsequently allowed the separation constant to have a small deviation from the above expression, which was explicitly frequency dependent. This allowed a nonzero intensity at the surface, but at the cost of rendering the separation of variables assumption invalid. Despite this treatment, a calculation of the surface flux from their Eq. 13 compared with the total emerging flux density at the surface of the sphere (Eq. C17) yields a $J=2H$ surface boundary condition. Under the two-stream approximation, a surface boundary condition $J=\sqrt{3}H$ is necessary to properly normalize the radiation pressure.

A different solution method is attempted here: namely, a continuous Fourier expansion in the frequency variable $\sigma$. The full solution is written
\be
J(r,\nu) & = & J_0(r,\nu) + J_{\rm bc}(r,\nu)
\ee
and
\be \label{eq:totalflux}
H(r,\nu) & = & H_0(r,\nu) + H_{\rm bc}(r,\nu).
\ee
Here $J_0$ and $H_0$ are the solutions of the inhomogeneous equation including the delta function source given in Equations \ref{eq:J0} and \ref{eq:H0}. The additional term $J_{\rm bc}$ must then be a solution of the homogeneous equation
\be
\frac{\partial^2J}{\partial r^2} + \frac{2}{r} \frac{\partial J}{\partial r}
+ \left( \frac{k}{\Delta} \right)^2 \frac{\partial^2 J}{\partial \sigma^2} &= & 0
\ee
with no delta function source term, and it must allow the boundary conditions to be satisfied at the surface. Since $J_0(R,
\nu)=0$, the surface boundary condition becomes
\be
J_{\rm bc}(R,\nu) - \sqrt{3} H_{\rm bc}(R,\nu) & = & 
J_{0}(R,\nu) + \frac{1}{\sqrt{3}k\phi} \frac{\partial J_{0}(R,\nu)}{\partial r} = 
\sqrt{3} H_0(R,\nu).
\label{eq:bc2}
\ee
Plugging in a frequency dependence $J \propto e^{is\sigma}$, for ``wavenumber" $s$, gives the equation for modified spherical Bessel functions of the first kind, $i_0(z)=\sinh(z)/z$.

% Fourier analysis stuff which we can take out
%\ifx
The solution can then be represented as
\be
J_{\rm bc}(r,\nu) & = & 
\int_{-\infty}^\infty \frac{ds}{2\pi} e^{is\sigma} A(s) 
\frac{i_0(krs/\Delta)}{i_0(kRs/\Delta)},
\label{eq:Jbc}
\ee
where $A(s)$ is the Fourier amplitude.

Plugging Eq. \ref{eq:Jbc} into Eq. \ref{eq:bc2} leads to the following equation for the Fourier amplitudes
\be
\int_{-\infty}^\infty \frac{ds}{2\pi} e^{is\sigma} A(s)
\left[ 1 + \left( \frac{s}{\sqrt{3} \Delta \phi} \right) \left( \frac{i_0^\prime(kRs/\Delta)}{i_0(kRs/\Delta)} \right) \right]
& = & \sqrt{3} H_0(R,\nu)
\label{eq:bc3}
\ee
where $H_0(R,\nu)$ is given by Eq. \ref{eq:H0surf}. Discretization of Eq. \ref{eq:bc3} for frequency variables $\sigma_i$ and wavenumbers $s_j$
leads to a set of coupled linear equations for the $A(s_j)$. We use equally-spaced points $\Delta \sigma = 2\sigma_{\rm max}/(N-1)$ and $\Delta s = 2\pi/(N\Delta \sigma)$, where $N$ is the number of points for each grid. The maximum frequency is set as $\sigma_{\rm max} = {\rm constant} \times \tau_0$, for a large enough constant that  the end of the frequency grid is at such small intensities that it does not affect the solution except close to the boundaries. \todo{The number of points was increased until the solution was well-resolved near line center, and only became inaccurate near the boundaries. (Is this still accurate?)} Given the Fourier amplitudes, $J_{\rm bc}$ is computed using Eq. \ref{eq:Jbc}, and the flux is given by
\be
H_{\rm bc}(r,\nu) & = & \left( \frac{-1}{3k\phi} \right)
\frac{\partial J_{\rm bc}(r,\nu)}{\partial r}
= \left( \frac{-1}{3k\phi} \right)
\int_{-\infty}^\infty \frac{ds}{2\pi} e^{is\sigma} A(s) 
\left( \frac{ks}{\Delta} \right) 
\left( \frac{i_0^\prime(krs/\Delta)}{i_0(kRs/\Delta)} \right).
\label{eq:Hbc}
\ee
The Bessel functions are finite at the center and rise steeply toward the surface when $kRs/\Delta \gg 1$.

@@

\begin{figure}
    \centering
    \includegraphics{taubc.pdf}
    \caption{Boundary condition term $H_{\rm bc}$ scaled by $(a\tau_0)^{1/3}$ shown for increasing line center optical depths. The $H_0$ solution minus the Monte Carlo data should be close to $H_{\rm bc}$ according to Eq. \ref{eq:totalflux}. As $\tau_0$ becomes larger, the agreement between these two improves, indicating that at large optical depths the relative size of the corrective boundary condition term scales as $(a\tau_0)^{-1/3}$}.
    \label{fig:taubc}
\end{figure}

@@

\todo{Brief description of MC code, citing Chenliang's paper. Discuss how MC works briefly, reference some papers, talk about problem setup specific to this work. Basically want to show that the MC approach in this paper is not new and doesn't need to be justified extensively.}

We now compare each of these solutions for surface flux with Monte Carlo simulations for a limited range of physical parameters $\tau_0$ and $x_s$. In Fig. \ref{fig:sol_mc_residual_0}, the Monte Carlo is shown with solutions $H_{\rm d}$, $H_{\rm 0}$, and $H_{\rm 0 + bc} = H_{\rm 0} + H_{\rm bc}$ overplotted for an optical depth of $\tau_0 = 10^7$ and photons emitted at line center $\sigma_s = \sigma_0$. The solution with the correct frequency-dependent boundary condition enforced, $H_{\rm 0 + bc}$, has considerably lower residuals to Monte Carlo than the other solutions, especially in the line wing. It is evident that the divergent solution $H_{\rm d}$ fails in the line wings. The $H_{\rm bc}$ term, when added to the solution $H_0$ derived by \citet{2006ApJ...649...14D}, adequately corrects for the apparent excess of flux in the peaks of the spectrum. This peak excess is also present in previous work that used the \citet{1990ApJ...350..216N} solution, such as \citet{2015MNRAS.449.4336S}. The boundary term resolves the slight deficit of $H_{\rm0}$ in the line wings, further improving agreement with the numerical result.

@@

>>>>

The solution of the radiative transfer equation for resonant scattering of Lyman $\alpha$ photons for a uniform sphere of constant gas density is found in the limit of large line center optical depth. A monochromatic source of photons is assumed at the center of the sphere, and the intensity within the sphere and at the surface is found with the Eddington approximation for the angular dependence. The solution can be represented a sum of three terms: a simple, analytic solution of the inhomogeneous equation which diverges at the center of the sphere and at the emission frequency; a semi-analytic solution of the homogeneous equation which is required for the mean intensity $J=0$ at the surface; and finally, a solution of the homogeneous equation which enforces the correct, zero-ingoing intensity boundary conditions at the surface. It is the compact, analytic form as well as the piece accounting for the correct, frequency dependent boundary condition that is novel in this study. The analytic solution is compared to numerical solutions with the Monte-Carlo method, which are valid at arbitrary optical depth, and the deviations from exact solution are investigated.

@@

The outer layers of planetary atmospheres are central to a planet's evolution, since they can shelter the lower atmosphere from high energy radiation and regulate the escape of gas into space. In atmospheres with thick layers of atomic hydrogen, stellar Lyman Alpha (Ly$\alpha$) radiation is especially important in this process as the optical depth at line center is large. Ionization and subsequent recombination can generate \lya within the planet's atmosphere, dissociating molecules in lower layers and causing heating by collisional de-excitation. \lya can also excite H atoms to the 2p state, creating a population of Balmer-line absorbers that can be observed via transmission spectroscopy.

@@

Hubble Space Telescope (HST) observations with STIS have found large \lya transit depths around a handful of exoplanets, indicating an extended distribution of hydrogen: \citet{2003Natur.422..143V, 2012A&A...543L...4L, 2012A&A...547A..18E, 2015Natur.522..459E,  2017A&A...597A..26B, 2017A&A...599L...3B, 2017A&A...602A.106B, 2018A&A...620A.147B, 2019AJ....158...50W, 2019EPSC...13.1928L, 2020ApJ...888L..21G,2021arXiv210309864B}. These observations have revealed a population of atoms extending out to distances of order the planetary radius for several planets around bright, nearby stars, motivating a study of the physics of \lya interactions with H atoms surrounding a planet. The importance of the \lya line in the atmospheric dynamics of these systems necessitates a careful treatment of resonant scattering in order to construct accurate models of atmospheric heating and escape. However, the extreme optical depths at \lya line center introduce a computational barrier to simulating radiative transfer via Monte Carlo methods.

@@

A third class of methods are modified random walk methods \citep{1984JCoPh..54..508F,2009A&A...497..155M, 2010A&A...520A..70R}. In this case a spherical volume (usually the largest within a simulation cell) is chosen when the scattering mean-free-path is small compared with the zone size.  An outgoing photon is then randomly sampled on the surface of this outgoing sphere by drawing its properties from distributions in outgoing frequencies, directions and escape times, based on solutions to the diffusion equation within a uniform sphere. A method similar to this has been applied by \citet{2006ApJ...645..792T} to Lyman $\alpha$ transfer using the \cite{1990ApJ...350..216N} solution.

The need for accurate distributions from which to draw outgoing photon properties motivates this work; namely, an analytic solution to the steady-state problem of \lya resonant scattering in spherical geometry. In addition, a semi-analytic solution to time-dependent \lya diffusion through the same medium is explored to quantify how long the photons are expected to spend in the domain before escaping. These solutions will be shown to accurately represent the photon distributions in frequency and time-to-escape at the surface of a spherical volume of gas, highlighting their potential for use as a Monte Carlo accelerant in magnetohydrodynamic calculations of phenomena involving \lya resonance-line radiation.

@@

\citet{1973MNRAS.162...43H, 1974MNRAS.166..373H} demonstrated that resonance-line radiation transfer in low-density, optically-thick media can be described by the Poisson equation. However, the accuracy of their solution is limited by the treatment of the boundary conditions. In their eigenfunction expansion, the spatial and frequency components of the differential equation are separated, yet the spatial eigenvalues are defined as being explicitly frequency-dependent. The resulting solution for large optical depths would suggest the mean intensity at the boundary of the slab is zero, which is unphysical. A more detailed discussion of the boundary condition is included in Section \ref{sec:steadystate}. \citet{1990ApJ...350..216N} follows this approach in an attempt to extend the result to media of intermediate optical depth, including the effects of scattering in the Doppler core of the line, but again does not use the correct boundary condition. \citet{?} improved on this approach, producing a divergent, inhomogeneous solution to the problem in 1D slab geometry that includes the effect of photon absorption within the medium. \citet{2006ApJ...649...14D} generalize the problem to spherical geometry, which is the same physical set up used here. Following this, \citet{2020arXiv200509692L} presents closed-form solutions for the radiation field in spherical geometry, utilizing the ``core skipping'' method discussed in \citet{2015MNRAS.449.4336S} to handle photon frequency redistribution within the Doppler core. This technique involves drawing the atom's velocity from a truncated Gaussian distribution, forcing photons back into the wing when they scatter into the core. This is motivated by the assumption that all core scattering results in zero spatial diffusion until the photon is redistributed back out into the wing where the mean free path becomes much larger and it can diffuse in space again. However, it is not clear whether this prescription is fully independent of the choice of velocity distribution.

@@

\subsection{Derivation of the Steady-State Solution}
Now, consider a sphere of radius $R$ and uniform density, with line-center optical depth $\tau_0 = (k/\sqrt{\pi}\Delta)R$. After changing variables from $\nu$ to $\sigma$ using $d\sigma = \sqrt{2/3}d\nu/(\Delta^2 \phi)$ in Eq. \ref{eq:rteqn} and performing all angular integrals, we obtain Eq. \ref{eq:finaleqn} (see Appendix \ref{app:rteqn_derivation}) and set $p=\alpha_{\rm abs}=0$ to eliminate photon destruction terms. With a photon emission term given by Eq. \ref{eq:jem}, this gives the transfer equation
\be
\nabla^2 J + \left( \frac{k}{\Delta} \right)^2 \frac{\partial^2 J}{\partial \sigma^2} & = & 
- \frac{ \sqrt{6} kL}{4\pi \Delta^2} \delta^3(\vec{x} - \vec{x}_s) \delta (\sigma - \sigma_{\rm s})
\label{eq:rt_no_destr}
\ee
where the form of the source term is in Eq. \ref{eq:jem_v2}. The boundary condition is that there is no incoming flux at the surface, which can be written
\citep{1986rpa..book.....R}
\be
J & = & \sqrt{3} H
\label{eq:bc}
\ee
at $r=R$. First consider the divergent solution. In terms of the dimensionless variables $x_1 = k(x-x_{\rm s})/\Delta$, $x_2=k(y-y_{\rm s})/\Delta$, $x_3=k(z-z_{\rm s})/\Delta$ and $x_4=\sigma - \sigma_{\rm s}$, this equation can be written as a Laplacian in the 4-dimensional space as
\be
\nabla_4^2 J & = & - \frac{\sqrt{6} k^2 L}{4\pi \Delta^3} \delta^4(x).
\ee
In terms of the 4-dimensional radius $r_4 = \sqrt{ (k/\Delta)^2\left[ (x-x_{\rm s})^2 + (y-y_{\rm s})^2 + (z-z_{\rm s})^2 \right] + (\sigma - \sigma_{\rm s})^2 }$, and assuming spherical symmetry about the source point, the equation becomes
\be
\frac{1}{r_4^3} \frac{\partial}{\partial r_4} \left( r_4^3 \frac{\partial J}{\partial r_4} \right) & = & - \frac{\sqrt{6} k^2 L}{4\pi \Delta^3} 
\frac{\delta(r_4)}{2\pi^2 r_4^3},
\ee
where the forms of the 4-dimensional Laplacian and delta function have been inserted. The ``divergent" solution due to the delta function is then
\be
J_{\rm d} & = & 
\frac{(\sqrt{6}k^2L)/(16\pi^3 \Delta^3)}{ (k/\Delta)^2\left[ (x-x_{\rm s})^2 + (y-y_{\rm s})^2 + (z-z_{\rm s})^2 \right] + (\sigma - \sigma_{\rm s})^2}.
\label{eq:Jd}
\ee
While this solution does not satisfy the surface boundary condition (Eq. \ref{eq:bc}), it is accurate near the delta-function source.
This solution generalizes the particular solution for slab geometry found by \citet{1990ApJ...350..216N}. Henceforth we will focus on the source at the center of the sphere, although we will retain the emission frequency at $\sigma=\sigma_{\rm s}$. The radial flux corresponding to Eq. \ref{eq:Jd} is
\be
H_{\rm d} & = & - \frac{1}{3k\phi} \frac{\partial J_d}{\partial r}
=  \left( \frac{1}{3k\phi} \right) 
\left( \frac{ \sqrt{6}k^3L }{ 8\pi^3 \Delta^4} \right)
\left( \frac{k r/\Delta}{ \left[ (kr/\Delta)^2 + (\sigma-\sigma_{\rm s})^2 \right]^2 } \right).
\label{eq:Hd}
\ee
It will be shown that Eq. \ref{eq:Hd} is a good approximation at $r=R$ near line center. However Eq. \ref{eq:Jd} is not a good approximation to the true solution, as it is too large at $r=R$ by a factor of $J_{\rm d}(R,\sigma)/ H_{\rm d}(R,\sigma) \sim a\tau_0/x^2 \sim (a\tau_0)^{1/3} \gg 1$.

@@

In the limit of large optical depth $B$, \citealt{1973MNRAS.162...43H} approximates the eigenvalues as $\lambda_n B \simeq \pi (n-1/2)$, which gives $J=0$ at the surface. This is essentially the result in Eq. \ref{eq:J0} above. However, since this solution gives zero at the surface, their Eq. 34 subsequently allowed the separation constant to have a small deviation from the above expression, which was explicitly frequency dependent. This allowed a nonzero intensity at the surface, but at the cost of rendering the separation of variables assumption invalid. Despite this treatment, a calculation of the surface flux from their Eq. 13 compared with the total emerging flux density at the surface of the sphere (Eq. C17) yields a $J=2H$ surface boundary condition. However, under the two-stream approximation, a surface boundary condition $J=\sqrt{3}H$ is necessary to properly normalize the radiation pressure.

A different solution method is attempted here: namely, a continuous Fourier expansion in the frequency variable $\sigma$. The full solution is written
\be
J(r,\sigma) & = & J_0(r,\sigma) + J_{\rm bc}(r,\sigma)
\ee
and
\be \label{eq:totalflux}
H(r,\sigma) & = & H_0(r,\sigma) + H_{\rm bc}(r,\sigma).
\ee
Here $J_0$ and $H_0$ are the solutions of the inhomogeneous equation including the delta function source given in Equations \ref{eq:J0} and \ref{eq:H0}. The additional term $J_{\rm bc}$ must then be a solution of the homogeneous equation
\be
\frac{\partial^2J}{\partial r^2} + \frac{2}{r} \frac{\partial J}{\partial r}
+ \left( \frac{k}{\Delta} \right)^2 \frac{\partial^2 J}{\partial \sigma^2} &= & 0
\ee
with no delta function source term, and it must allow the boundary conditions to be satisfied at the surface. Since $J_0(R,
\nu)=0$, the surface boundary condition becomes
\be
J_{\rm bc}(R,\sigma) - \sqrt{3} H_{\rm bc}(R,\sigma) & = & 
J_{0}(R,\sigma) + \frac{1}{\sqrt{3}k\phi} \frac{\partial J_{0}(R,\sigma)}{\partial r} = 
\sqrt{3} H_0(R,\sigma).
\label{eq:bc2}
\ee
Plugging in a frequency dependence $J \propto e^{is\sigma}$, for ``wavenumber" $s$, gives the equation for modified spherical Bessel functions of the first kind, $i_0(z)=\sinh(z)/z$.

% Fourier analysis stuff which we can take out
%\ifx
The solution can then be represented as
\be
J_{\rm bc}(r,\sigma) & = & 
\int_{-\infty}^\infty \frac{ds}{2\pi} e^{is\sigma} A(s) 
\frac{i_0(krs/\Delta)}{i_0(kRs/\Delta)},
\label{eq:Jbc}
\ee
where $A(s)$ is the Fourier amplitude.

Plugging Eq. \ref{eq:Jbc} into Eq. \ref{eq:bc2} leads to the following equation for the Fourier amplitudes
\be
\int_{-\infty}^\infty \frac{ds}{2\pi} e^{is\sigma} A(s)
\left[ 1 + \left( \frac{s}{\sqrt{3} \Delta \phi} \right) \left( \frac{i_0^\prime(kRs/\Delta)}{i_0(kRs/\Delta)} \right) \right]
& = & \sqrt{3} H_0(R,\sigma)
\label{eq:bc3}
\ee
where $H_0(R,\sigma)$ is given by Eq. \ref{eq:H0surf}. Discretization of Eq. \ref{eq:bc3} for frequency variables $\sigma_i$ and wavenumbers $s_j$
leads to a set of coupled linear equations for the $A(s_j)$. We use equally-spaced points $\Delta \sigma = 2\sigma_{\rm max}/(N-1)$ and $\Delta s = 2\pi/(N\Delta \sigma)$, where $N$ is the number of points for each grid. The maximum frequency is set as $\sigma_{\rm max} = {\rm constant} \times \tau_0$, for a large enough constant that  the end of the frequency grid is at such small intensities that it does not affect the solution except close to the boundaries. \todo{The number of points was increased until the solution was well-resolved near line center, and only became inaccurate near the boundaries. (Is this still accurate?)} Given the Fourier amplitudes, $J_{\rm bc}$ is computed using Eq. \ref{eq:Jbc}, and the flux is given by
\be
H_{\rm bc}(r,\sigma) & = & \left( \frac{-1}{3k\phi} \right)
\frac{\partial J_{\rm bc}(r,\sigma)}{\partial r}
= \left( \frac{-1}{3k\phi} \right)
\int_{-\infty}^\infty \frac{ds}{2\pi} e^{is\sigma} A(s) 
\left( \frac{ks}{\Delta} \right) 
\left( \frac{i_0^\prime(krs/\Delta)}{i_0(kRs/\Delta)} \right).
\label{eq:Hbc}
\ee
The Bessel functions are finite at the center and rise steeply toward the surface when $kRs/\Delta \gg 1$.

@@

\begin{figure}
    \centering
    \includegraphics{taubc.pdf}
    \caption{Boundary condition term $H_{\rm bc}$ scaled by $(a\tau_0)^{1/3}$ shown for increasing line center optical depths. The Monte Carlo data minus the $H_0$ solution should be close to $H_{\rm bc}$ according to Eq. \ref{eq:totalflux}. As $\tau_0$ becomes larger, the agreement between these two improves, indicating that at large optical depths the relative size of the corrective boundary condition term scales as $(a\tau_0)^{-1/3}$}.
    \label{fig:taubc}
\end{figure}

@@

%Brief description of MC code, citing Chenliang's paper. Discuss how MC works briefly, reference some papers, talk about problem setup specific to this work. Basically want to show that the MC approach in this paper is not new and doesn't need to be justified extensively.

The Monte Carlo method is used to solve the transfer equation numerically as a point of reference for these analytic solutions. Photon packets are initialized at a monochromatic source frequency $x_s$ and are allowed to propagate through the medium until escaping, at which point their positions, outgoing angles, and frequencies are logged. Frequency redistribution is calculated at each scattering, including the effects of recoil, to obtain the spectrum at the surface of the spherical simulation domain. Further details of the Monte Carlo simulation are discussed in \cite{2017ApJ...851..150H}.

@@

We now compare each of the solutions for surface flux with Monte Carlo simulations for a limited range of physical parameters $\tau_0$ and $x_s$. In Fig. \ref{fig:sol_mc_residual_0}, the Monte Carlo is shown with solutions $H_{\rm d}$, $H_{\rm 0}$, and $H_{\rm 0 + bc} = H_{\rm 0} + H_{\rm bc}$ overplotted for an optical depth of $\tau_0 = 10^7$ and photons emitted at line center $\sigma_s = \sigma_0$.

The solution with the correct frequency-dependent boundary condition enforced, $H_{\rm 0 + bc}$, has considerably lower residuals to Monte Carlo than the other solutions, especially in the line wing. It is evident that the divergent solution $H_{\rm d}$ fails in the line wings. The $H_{\rm bc}$ term, when added to the solution $H_0$ derived by \citet{2006ApJ...649...14D}, adequately corrects for the apparent excess of flux in the peaks of the spectrum. This peak excess is also present in previous work that used the \citet{1990ApJ...350..216N} solution, such as \citet{2015MNRAS.449.4336S}. The boundary term resolves the slight deficit of $H_{\rm0}$ in the line wings, further improving agreement with the numerical result.
